Rising from the Ashes: How the Milky Way Got Its Scars

### 2024-07-01 paper/src/abs.tex
First draft edits

\begin{abstract}
    The nuclear abundance distribution of stars encodes the history of the gas-phase abundance in the Milky Way. Without a large, unbiased sample of stellar ages, the exact timing and nature of this history must be \textit{inferred} from the abundances. In the two-dimensional plane of \alphaFe-\FeH, it is now clear that two separate populations exist - the low-$\alpha$ and high-$\alpha$ sequences. Structure in the nuclear abundance distribution can arise from many processes -- proposals include specific gas infall scenarios, radial migration, high-redshift clump formation, and various effects associated with galaxy mergers, among others. In this work, we demonstrate another possible avenue for structure formation with clear observational predictions. We propose that the Galaxy underwent a starburst followed by a brief (hundreds of Myr) quiescent phase -- i.e., the Galaxy underwent a post-starburst rejuvenation sequence at $z\sim2$. A natural consequence of the quiescent phase is that stars in the valley of the bimodality do not form because: (1) the absence of enrichment from high-mass stars leads to a rapid reduction in \alphaFe{}, and (2) any time the gas spends in the abundance valley is deemphasized in the present day distribution because the star formation rate is lower. With a set of idealized merger simulations, we demonstrate the feasibility of this proposal. This ``phoenix hypothesis'' predicts a $\sim300\,\Myr$ gap in stellar ages at a fixed \FeH{} and that stars which form directly after this gap would have lower \alphaFe{} than stars which form slightly ($\sim1\,\Gyr$) later.
  \end{abstract}

>>>>

\begin{abstract}
    The nuclear abundance distribution of stars encodes the history of the gas-phase abundance in the Milky Way. Without a large, unbiased sample of highly precise stellar ages, the exact timing and nature of this history must be \textit{inferred} from the abundances. In the two-dimensional plane of \alphaFe-\FeH, it is now clear that two separate populations exist -- the low-$\alpha$ and high-$\alpha$ sequences. Structure in the nuclear abundance distribution can arise from many processes -- proposals include specific gas infall scenarios, radial migration, high-redshift clump formation, and various effects associated with galaxy mergers, among others. In this work, we demonstrate another possible avenue for structure formation with clear observational predictions. We propose that the Galaxy underwent a starburst followed by a brief (hundreds of Myr) quiescent phase -- i.e., the Galaxy underwent a post-starburst rejuvenation sequence at $z\sim2$. A natural consequence of the quiescent phase is that stars in the valley of the bimodality do not form because: (1) the absence of enrichment from high-mass stars leads to a rapid reduction in \alphaFe{}, and (2) any time the gas spends in the abundance valley is deemphasized in the present day distribution because the star formation rate is lower. With a set of idealized merger simulations, we demonstrate the feasibility of this proposal. This ``phoenix hypothesis'' predicts a $\sim300\,\Myr$ gap in stellar ages at a fixed \FeH{} and that stars which form directly after this gap would have lower \alphaFe{} than stars which form slightly ($\sim1\,\Gyr$) later.
  \end{abstract}

### 2024-07-01 paper/src/con.tex
First draft edits

In this work, using idealized merger simulations, we have shown that (1) a starburst, followed by (2) a brief ($\sim300\,\Myr$) quiescent period, followed by (3) rejuvenation can lead to a bimodal abundance structure. We found that the starburst-to-quiescent period allowed for prodigious Fe-enrichment without any corresponding $\alpha$-enrichment. This ``phoenix hypothesis'' would predict that, for stars at a given \FeH{}, there would be a $\sim300\,\Myr$ gap in stellar ages at $\sim8\,\Gyr$. Furthermore, stars which form directly after this gap would have lower \alphaFe{} than stars which form slightly later ($\sim1\,\Gyr$).

The bimodality in the \alphaFe{}-\FeH{} plane of stellar abundances points to deeper, unobserved structure. That structure is a record of the complicated processes that gave rise to the Galaxy. In the absence of highly precise stellar ages, that history must be largely inferred.

>>>>

In this work, using idealized merger simulations, we have shown that (1) a starburst, followed by (2) a brief ($\sim300\,\Myr$) quiescent period, followed by (3) rejuvenation can lead to a bimodal abundance structure. We found that the starburst-to-quiescent period allowed for prodigious Fe-enrichment without any contemperaneous $\alpha$-enrichment. This sequence of events, somewhat resembling the life cycle of a phoenix, would predict that for stars at a given \FeH{}, there would be a $\sim300\,\Myr$ gap in stellar ages at $\sim8\,\Gyr$. Furthermore, stars which form directly after this gap would have lower \alphaFe{} than stars which form slightly later ($\sim1\,\Gyr$).

The bimodality in the \alphaFe{}-\FeH{} plane of stellar abundances points to deeper, unobserved structure. That structure is a record of the complicated processes that gave rise to the Galaxy. In the absence of highly precise stellar ages, that history must be largely inferred. There are many ways in which a galaxy could end up with the coarse-grained structure we can observe in the present day. We have argued in this work that a brief, starburst-driven quiescent phase is one of them.

### 2024-07-01 paper/src/dis.tex
First draft edits

This drop can be understood in terms of the mechanics of the galaxy formation model as described in Section~\ref{ssec:gfm}. When a star particle forms, it only takes $\sim40\,\Myr$ before the last Type II SN has gone off and Fe enrichment is dominated by Type Ia SNe. As a result, during the quiescent phase there is relatively little $\alpha$-element production. On the other hand, star particles are still able to generate Fe through Type Ia SNe which have a much longer period of influence. In the TNG model, the number of Type Ia SNe decays with time as $(t/\tau_8)^{-s}$, where $\tau_8=40\,\Myr$ is the lifetime of an $8\,\Msun$ star and $s=1.12$. Therefore, the burst of star formation at $t\sim2.25\,\Gyr$ is able to generously produce Fe starting at $2.5\,\Gyr$, when the drop in \MgFe{} is the most steep.

@@

\citet{2018MNRAS.477.5072M} found that Milky Way-like bimodalities are rare in EAGLE, occuring in $\sim5\%$ of galaxies. \citet{2021MNRAS.501..236D,2022MNRAS.515.1430D} showed that merger-induced quenching in zooms can occur in the EAGLE model \citep[see also][]{2017MNRAS.465..547P}. However, the situation may be different in the lower resolution large box. Furthermore, if the proposed starburst-quenching phase is driven by AGN feedback, then the outcome of any particular cosmological simulation with regards to the bimodality is intimately tied to its AGN model. Unfortunately, such models are highly uncertain \red{cite}.

\citet{2023arXiv231016085K} explored the impact of quenching in the Magneticum Pathfinder suite. They found that galaxies which quench undergo a starburst followed by an AGN-driven quenching phase. In the post-starburst regime, they claim galaxies are $\alpha$-enhanced. From their Figure~8 it is difficult to quantify the difference between the quenched and non-quenched galaxies as the two populations almost completely overlap. We do find that the bulk stellar \MgFe{} is enhanced after the merger in our fiducial simulation, but only at the $\sim0.01\,\dex$ level.

\subsection{Infall Interpretations}\label{ssec:dilute}
In some previous work, it was reported that the bimodality is a consequence of a sudden deposition of metal-poor, $\alpha$-poor gas by a satellite or cosmological filament -- i.e., a ``dilution'' \citep{2020MNRAS.491.5435B,2021MNRAS.503.5846R}. The separation of the sequences follows from the rapidity of the dilution. This was elaborated upon by \citet{2021MNRAS.503.5868R} who described a setup where the low-$\alpha$ disk forms inclined relative to the high-$\alpha$ disk.

@@

The longer-standing two-infall class of models argue that the two sequences diverge due to two episodic accretion episodes, with some possible enrichment of the second episode arising from an associated satellite \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}. As mentioned in the previous paragraph, our simulations do not support a strict gas dilution interpretation, but it was recently elaborated by \citet{2024arXiv240511025S} that these models also argue for a star formation gap between the two accretion episodes \citep[see also][]{1996ASPC...92..307G,1998A&A...338..161F,2000A&A...358..671G}. This gap is starvation-driven and can last several \Gyr{}. The present work argues for a starburst-driven quiescence followed by a rapid rejuvenation, with the entire process taking less than $1\,\Gyr$ and the gap only lasting a few hundred \Myr{}. The physical origin and some details are different, but one can appreciate that the bimodality arises from the same physical process.\footnote{Compare Section~\ref{ssec:formqui} to the first key result in the Conclusions of \citet{2024arXiv240511025S}.}

\subsection{Observational Connections: High Redshift Quenching}\label{ssec:obshiz}
Our main argument in this paper is that the Milky Way underwent a brief quiescent phase induced by a starburst at $z\sim2$. In observations, a Milky Way-analogue would then be identified as a post-starburst galaxy (PSB). With abundance matching, we expect the Milky Way's total stellar mass to be $\sim10^{10.3}\,\Msun$ at $z\sim2$ \citep{2013ApJ...771L..35V}. A number of authors have explored PSBs and quiescent galaxies at slightly higher masses at $z\sim2$, with large advances in the post-JWST era.

First, PSBs are not uncommon. \citet{2023ApJ...953..119P} found that in massive galaxies ($M_* > 10^{10.6}\,\Msun$) the fraction of PSBs increases from $\sim2.7\%$ ($99/3655$) at $1.0 < z < 1.44$ to $\sim8\%$ ($89/1118$) at $2.16 < z < 2.5$. Later, \citet{2024arXiv240417945P} found that $\sim10\%$ of galaxies at $\sim10^{10.3}\,\Msun$ are quenched \citep[consistent with][]{2013ApJ...777...18M}, and $\sim30\%$ of their total sample is a PSB at $z\sim2$. If these galaxies can be quickly rejuvenated, as the system studied in this work would suggest, then the total fraction of galaxies that go through a starburst-quenching phase may be higher.

Furthermore, \citet{2023arXiv231215012C} found that lower mass quiescent galaxies (towards $10^{10.3}\,\Msun$) tend to be younger and more disky, pointing to a merger driven scenario. \citet{2023arXiv231212207A} found that at even lower masses ($<10^{9.5}\,\Msun$) Furthermore, the AGN quenching mechanism, which we believe to be responsible for the quenching in our system (Appendix~\ref{app:cause_qui}), is thought to operate at these redshifts \citep[e.g.][and references therein]{2023arXiv230805795B,2024arXiv240417945P,2024arXiv240518685M}.

@@

And so, this test may only be possible to a reader in the distant future. However, the situation might not be completely hopeless. The necessary sample size is small, and so copious time and instruments can be dedicated. Moreover, only \textit{relative} ages need to be determined, so a sort of differential approach may yield interesting results. In addition to the presence of the gap, our work also predicts that the timing of the gap is \FeH{} dependent.

@@

There is also the perennial problem of diffusion within the hydrodynamics solver. In purely Lagrangian solvers, there is no diffusion between resolution elements, while in Eulerian codes the diffusion can be quite high. AREPO limits the numerical diffusion by allowing the mesh to move in a quasi-Lagrangian manner, and using a second-order solver \citep{2010MNRAS.401..791S}. In FIRE-2 \citep{2018MNRAS.480..800H}, which uses the Lagrangian code GIZMO \citep{2015MNRAS.450...53H}, a subgrid turbulent metal diffusion model was used. It would be interesting to see how models with different diffusivity properties would relax or strengthen the necessity of a quiescent period to produce a bimodality.

>>>>

This drop can be understood in terms of the mechanics of the galaxy formation model as described in Section~\ref{ssec:gfm}. When a star particle forms, it only takes $\sim40\,\Myr$ before the last Type II SN has gone off and its material enrichment is dominated by Fe from Type Ia SNe. As a result, during the quiescent phase which lasts a few hundred Myr, there is relatively little $\alpha$-element production. On the other hand, star particles are still able to generate Fe through Type Ia SNe which have a longer period of influence. The fact that the quiescent period is preceded by a starburst is important. Even though star particles produce Fe through the Type Ia channel for Gyr, younger star particles still produce much more Fe than older star particles. In the TNG model, the number of Type Ia SNe decays with time as $(t/\tau_8)^{-s}$, where $\tau_8=40\,\Myr$ is the lifetime of an $8\,\Msun$ star and $s=1.12$. Therefore, the burst of star formation at $t\sim2.25\,\Gyr$ is able to generously produce Fe starting at $2.5\,\Gyr$, when is indeed when the drop in \MgFe{} is the most steep.

@@

\citet{2018MNRAS.477.5072M} found that Milky Way-like bimodalities are rare in EAGLE, occuring in $\sim5\%$ of galaxies. \citet{2021MNRAS.501..236D,2022MNRAS.515.1430D} showed that merger-induced quenching in zooms can occur in the EAGLE model \citep[see also][]{2017MNRAS.465..547P}. However, the situation may be different in the lower resolution large box. Furthermore, if the proposed starburst-quenching phase is driven by AGN feedback, then the outcome of any particular cosmological simulation with regards to the bimodality is intimately tied to its AGN model. Unfortunately, such models are highly uncertain \citep[e.g.][]{2022MNRAS.511.3751H}.

\citet{2023arXiv231016085K} explored the impact of quenching in the Magneticum Pathfinder suite. They found that galaxies which quench undergo a starburst followed by an AGN-driven quenching phase. In the post-starburst regime, they claim galaxies are $\alpha$-enhanced. We do find that the bulk stellar \MgFe{} is enhanced after the merger in our bimodal simulation compared to the isolated simulation, but only at the $\sim0.01\,\dex$ level.

\subsection{Infall Interpretations}\label{ssec:dilute}
In some previous work, it was reported that the bimodality is a consequence of a sudden deposition of metal-poor, $\alpha$-poor gas by a satellite or cosmological filament -- i.e., a ``dilution'' \citep{2020MNRAS.491.5435B,2021MNRAS.503.5846R}. The separation of the sequences follows from the rapidity of the dilution. This was elaborated upon by \citet{2021MNRAS.503.5868R} who described a zoom where the low-$\alpha$ disk forms out of a relatively pristine cosmological filament. This disk is inclined relative to the high-$\alpha$ disk, with the two disks later tidally realigning. The longer-standing two-infall class of models argue that the two sequences diverge due to two episodic accretion episodes, with some possible enrichment of the second episode arising from an associated satellite \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}.

@@

It was recently elaborated by \citet{2024arXiv240511025S} that these models also argue for a star formation gap between the two accretion episodes \citep[see also][]{1996ASPC...92..307G,1998A&A...338..161F,2000A&A...358..671G,2015A&A...578A..87S,2020A&A...640A..81N}. This gap is starvation-driven and can last several \Gyr{}. The present work argues for a starburst-driven quiescence followed by a rapid rejuvenation, with the entire process taking less than $1\,\Gyr$ and the gap only lasting a few hundred \Myr{}. The physical origin and some details are different, but one can appreciate that the bimodality arises from the same physical process.\footnote{Compare Section~\ref{ssec:formqui} to the first key result in the Conclusions of \citet{2024arXiv240511025S}.}

\subsection{Observational Connections: High Redshift Quenching}\label{ssec:obshiz}
If the Milky Way underwent a starburst to quiescent to rejuvenation sequence, as we argue in this paper, then an an analogue at $z\sim2$ would be identified as a post-starburst galaxy (PSB). With abundance matching, we expect the Milky Way's total stellar mass to be $\sim10^{10.3}\,\Msun$ at $z\sim2$ \citep{2013ApJ...771L..35V}. A number of authors have explored PSBs and quiescent galaxies at slightly higher masses at $z\sim2$, with large advances in the post-JWST era.

First, PSBs are not uncommon. \citet{2023ApJ...953..119P} found that in massive galaxies ($M_* > 10^{10.6}\,\Msun$) the fraction of PSBs (inferred ages $< 800\,\Myr$) increases from $\sim2.7\%$ ($99/3655$) at $1.0 < z < 1.44$ to $\sim8\%$ ($89/1118$) at $2.16 < z < 2.5$ \citep[see also][]{2012ApJ...745..179W,2019ApJ...874...17B}. Later, \citet{2024arXiv240417945P} found that $\sim10\%$ of galaxies at $\sim10^{10.3}\,\Msun$ are quenched \citep[consistent with][]{2013ApJ...777...18M}, and $\sim30\%$ of their quiescent sample is a PSB at $z\sim2$. If these galaxies can be quickly rejuvenated, as the system studied in this work would suggest, then the total fraction of galaxies that go through a starburst-quenching phase may be higher.

Furthermore, \citet{2023arXiv231215012C} found that lower mass quiescent galaxies (towards $10^{10.3}\,\Msun$) tend to be younger and more disky, pointing to a merger driven scenario. \citet{2023arXiv231212207A} found that at even lower masses ($<10^{9.5}\,\Msun$) There is also evidence that the AGN quenching mechanism, which we believe to be responsible for the quenching in our system (Appendix~\ref{app:cause_qui}), is operating at these redshifts \citep[e.g.][and references therein]{2023arXiv230806317D,2024arXiv240417945P,2024arXiv240518685M,2024Natur.630...54B}.

Overall, the starburst-driven quenching mechanism proposed in this work is allowed by high-redshift observations.

@@

And so, this test may only be possible to a reader in the distant future. However, the situation might not be completely hopeless. The necessary sample size is small, and so copious time and instruments can be dedicated to each source. Moreover, only \textit{relative} ages need to be determined, so a sort of differential approach may be prudent.

In addition to the presence of the gap, our work also predicts that: (1) the timing of the gap is \FeH{} dependent, and that (2) stars which form immediately after the gap are $\alpha$-poor compared to stars which form slightly later (see Figure~\ref{fig:alpha_vs_tform}). However, both of these tests rely on the precise age estimates.

@@

There is also the perennial problem of diffusion within the hydrodynamics solver. In purely Lagrangian solvers, there is no diffusion between resolution elements, while in Eulerian codes the diffusion can be quite high.\footnote{No galaxy formation simulation is fully Lagrangian since there must be, at a minimum, mass exchange between star particles and gas. Here we just mean that there is no mass exchange between gas elements.} AREPO limits the numerical diffusion by allowing the mesh to move in a quasi-Lagrangian manner, and using a second-order solver \citep{2010MNRAS.401..791S}. In FIRE-2 \citep{2018MNRAS.480..800H}, which uses the Lagrangian code GIZMO \citep{2015MNRAS.450...53H}, a subgrid turbulent metal diffusion model was used. It would be interesting to see how models with different diffusivity properties would relax or strengthen the necessity of a quiescent period to produce a bimodality.

### 2024-07-01 paper/src/int.tex
Some Sarah changes
First draft edits

% Why alpha-elements are important
The distribution of elemental abundances is a high dimensional space \citep[e.g., 32 elements in][]{2024ApJ...961L..41J}. However, this space is highly degenerate, and so the effective number of dimensions is much smaller -- even possibly compressed to just \FeH{} and age \citep{2019ApJ...883..177N}. Two elements in particular have received particular interest - Fe and $\alpha$-elements. Type Ia and Type II supernovae are the main contributors of Fe and $\alpha$ enrichment. Fe is broadly produced in both types, and so its abundance is a proxy for the total metallicity of a star. On the other hand, $\alpha$-elements are mainly produced in Type II supernovae. The ratio of $\alpha$-elements to Fe (\alphaFe{}) is then a measure of the relative contributions of Type Ia and II SNe to the enrichment of a parcel of gas. It has therefore become common to compress the high-dimensional abundance space to the two dimensional \alphaFe{}-\FeH{} plane.

% Observed chemical bimodality in the Milky Way
It is now well-established that, in the Milky Way, a bimodality exists in the \alphaFe{}-\FeH{} plane \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R,2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}. This bimodality is typically separated into the high-$\alpha$ and low-$\alpha$ sequences. The high-$\alpha$ sequence is more centrally compact and vertically extended than the low-$\alpha$ sequence.

% Different explanations of bimodality
Naturally, many different processes that could lead to structure in the abundance plane have been discussed in the literature. They can be separated into two broad classes: (1) structure arising from internal processes, and (2) structure arising from external processes. We will summarize these classes of explanations in turn.

An early explanation of the bimodality is the two-phase gas infall model \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}. In this model, the thick disk first forms rapidly from an initial infall of gas. Because the typical SFR is high, these stars are $\alpha$-enhanced. In some variants, star formation halts completely before a second supply of pristine gas falls into the Galaxy. This dilutes the gas supply from which the thin disk forms more gradually. The thin disk is then more $\alpha$-poor because its associated SFR is lower.

@@

Yet another explanation invoking an internal process is that the formation of clumps at high redshift are responsible for both the chemistry and dynamics of the high-$\alpha$ sequence \citep{2019MNRAS.484.3476C,2020MNRAS.492.4716B,2021MNRAS.502..260B,2023ApJ...953..128G}. Instabilities are thought to form clumps in gas-rich disks, and such clumps are seen at intermediate redshifts \citep[$z\sim2$;][]{2005ApJ...627..632E,2007ApJ...658..763E}. These clumps then self-enrich, forming $\alpha$-enhanced stars. The high-$\alpha$ sequence stops forming once the gas fraction is low enough for the instabilities to no longer arise. This model predicts that the high-$\alpha$ and low-$\alpha$ sequences form simultaneously.

@@

In this work, we explored an idealized merger simulation which resembles the merger between the Milky Way and GSE. We found that in certain configurations, which have only minor differences in their orbit, a bimodal distribution of stellar abundances in the \alphaFe{}-\FeH{} plane is produced. This is shown in Figure~\ref{fig:fig1}. The configurations which produce a bimodality are associated with a starburst followed by a quiescent phase followed by rejuvenation $\sim300\,\Myr$ later.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{figure1.pdf}
  \caption{\textbf{The abundance bimodality seen in the Milky Way can be reproduced in some idealized merger simulations.} In the upper panels, we show the distribution of stars in the \MgFe-\FeH plane. The left panel shows the observed distribution in the Milky Way from ASPCAP DR17 \citep[][J.A.~Holtzman et al., in preparation]{2016AJ....151..144G}, while the right two panels show two idealized merger simulations. The idealized merger simulations are nearly identical, except that the bimodal simulation has a starting radius of $129\,\kpc$, while the unimodal simulation has a starting radius of $142\,\kpc$. We emphasize that the labels ``unimodal'' and ``bimodal'' are of the \textit{outcome} of the simulation, and do not reflect a particular choice in the setup. The bottom panels show the distribution of \MgFe at fixed \FeH. The colors indicate the fixed \FeH values, which are $-0.5$, $-0.375$, $-0.25$, and $0.25$. The Milky Way (left panels) exhibits a strong bimodal distribution of \MgFe at various \FeH. The idealized merger simulation marked as bimodal (center panels) also exhibits a bimodal distribution of \MgFe, though the structure is not as strongly defined. The idealized merger simulation marked as unimodal exhibits only weak structure, if any at all.}
  \label{fig:fig1}
\end{figure*}

>>>>

% Why alpha-elements are important
The distribution of elemental abundances is a high dimensional space \citep[e.g., 32 elements in][]{2024ApJ...961L..41J}. However, this space is highly degenerate, and so the effective number of dimensions is much smaller -- even possibly compressed to just \FeH{} and age \citep{2019ApJ...883..177N}. Two elements have received particular interest - Fe and elements produced by the $\alpha$-process. Type Ia and Type II supernovae are the main contributors of Fe and $\alpha$ enrichment. Fe is broadly produced in both types, and so its abundance is a proxy for the total metallicity of a star. On the other hand, $\alpha$-elements are mainly produced in Type II supernovae. The ratio of $\alpha$-elements to Fe (\alphaFe{}) is then a measure of the relative contributions of Type Ia and II SNe to the enrichment of a parcel of gas. It has therefore become common to compress the high-dimensional abundance space to the two dimensional \alphaFe{}-\FeH{} plane.

% Observed chemical bimodality in the Milky Way
It is now well-established that, in the Milky Way, a bimodality exists in the \alphaFe{}-\FeH{} plane \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R,2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}. This bimodality is typically separated into the high-$\alpha$ and low-$\alpha$ sequences. These sequences are also associated with differences in their physical extent -- the high-$\alpha$ sequence is more centrally compact and vertically extended than the low-$\alpha$ sequence.

% Different explanations of bimodality
Naturally, many different processes that could lead to structure in the abundance plane have been discussed in the literature. An early explanation of the bimodality is the two-phase gas infall model \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}. In this model, the thick disk first forms rapidly from an initial infall of gas. Because the typical SFR is high, these stars are $\alpha$-enhanced. In some variants, star formation halts completely before a second supply of pristine gas falls into the Galaxy \citep[][and references therein]{2024arXiv240511025S}. This dilutes the gas supply from which the thin disk forms more gradually. The thin disk is then more $\alpha$-poor because its associated SFR is lower.

A later argument by \citet{2021MNRAS.501.5176K} asserts that the two sequences follow from two phases of gas infall, except driven by stellar feedback instead of cosmological inflow. An initial bursty phase follows from the direct collapse of the gaseous halo. The disk has a high SFR leading to the formation of the high-$\alpha$ sequence. Feedback then halts the inflow, and a slower accretion of high-angular momentum and metal-rich gas commences, forming the low-$\alpha$ sequence.

@@

Yet another explanation, which also invokes an internal process, is that the formation of clumps at high redshift are responsible for both the chemistry and dynamics of the high-$\alpha$ sequence \citep{2019MNRAS.484.3476C,2020MNRAS.492.4716B,2021MNRAS.502..260B,2023ApJ...953..128G}. Instabilities are thought to form clumps in gas-rich disks, and such clumps are seen at intermediate redshifts \citep[$z\sim2$;][]{2005ApJ...627..632E,2007ApJ...658..763E}. These clumps then self-enrich, forming $\alpha$-enhanced stars. The high-$\alpha$ sequence stops forming once the gas fraction is low enough for the instabilities to no longer arise. This model predicts that the high-$\alpha$ and low-$\alpha$ sequences form simultaneously.

@@

In this work, we explored an idealized simulation which resembles the merger between the Milky Way and GSE. We found that in certain configurations, which have only minor differences in their orbit, a bimodal distribution of stellar abundances in the \alphaFe{}-\FeH{} plane is produced. This is shown in Figure~\ref{fig:fig1}. The configurations which produce a bimodality are associated with a starburst followed by a quiescent phase followed by rejuvenation $\sim300\,\Myr$ later.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{figure1.pdf}
  \caption{\textbf{The abundance bimodality seen in the Milky Way can be reproduced in some idealized merger simulations.} In the upper panels, we show the distribution of stars in the \MgFe{}-\FeH{} plane. The lower panels show the distribution of \MgFe{} at a fixed \FeH{} bin of width $0.05\,\dex$. The colors indicate the fixed \FeH values, which are $-0.5$, $-0.25$, $0$, and $0.25$. The left column shows the observed distribution in the Milky Way from ASPCAP DR17 \citep[][J.A.~Holtzman et al., in preparation]{2016AJ....151..144G}, while the right two columns show two idealized merger simulations. The idealized merger simulations are nearly identical, except that in the bimodal simulation the satellite has a starting radius of $129\,\kpc$, while in the unimodal simulation it has a starting radius of $142\,\kpc$. The labels ``unimodal'' and ``bimodal'' are of the \textit{outcome} of the simulation, and do not reflect a particular choice in the setup. The Milky Way (left column) exhibits a strong bimodal distribution of \MgFe{} at various \FeH{}. The idealized merger simulation marked as bimodal (center column) also exhibits a bimodal distribution of \MgFe{}, though the structure is not as strongly defined. The idealized merger simulation marked as unimodal (right column) exhibits only weak structure, if any at all.}
  \label{fig:fig1}
\end{figure*}
